---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Noah Booker"
date: "03/18/25"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
##library(dsbox) 
```
To install dsbox, it wanted me to update R. Instead I just downloaded the two datasets manually.
```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r inspecting_dennys_dataset}
load("~/Documents/Documents/Studies/School/WFU/DS4P/Lab_4/data/dennys.rda")
glimpse(dennys)
```

The dennys dataset has 1,643 observations of 6 variables. The observations represent Denny's restaurants and the variables represent various characteristics of each restaurant's location (address, city, state, zipcode, longitude, and latitude).


### Exercise 2
```{r inspecting_laquinta_dataset}
load("~/Documents/Documents/Studies/School/WFU/DS4P/Lab_4/data/laquinta.rda")
glimpse(laquinta)
```

The laquinta dataset has 909 observations of 6 variables. The observations represent La Quinta locations and the variables represent various characteristics of each hotel's location (address, city, state, zipcode, longitude, and latitude).


### Exercise 3
According to their website, there are La Quinta locations outside the US. There are locations in Canada, Mexico, China, New Zealand, Turkey, United Arab Emirates, Chile, Colombia, and Ecuador.
Denny's website only allows me to either search locations using a city, zipcode, or my current location or to browse US locations. So, I am ucertain whether there are Denny's locations outisde the US.

### Exercise 4
To examine the datasets to see whether La Quinta and Denny's have location outside of the US I could use something like dennys %>% distinct(state) for each dataset. Because they both have variables for the state of the location, I could use the distinct function to visually examine all the states that appear in the datasets to see if any seem like non-US states.

### Exercise 5
```{r nonUSdennys}
dennys %>%
  filter(!(state %in% states$abbreviation))
```
This filter gives a tibble with 0 rows, indicating that there are no Denny's locations in the dataset in states other than US states.

### Exercise 6
```{r add_country_var}
dennys <- dennys %>% 
  mutate(country = "United States")
```
Created a country variable for the dennys dataset.
Add exercise headings as needed.

